\textbf{Are regime changes always bad for markets? Evidence from daily financial data}

### 2021-07-01 doc/regime_changes_note.tex
Further note revisions/edits/cuts

\section{Reexamining how markets respond to irregular regime changes}

Conventional theory and cross-country empirical evidence suggests that political instability depresses financial returns \citep{irshad2017relationship, le2006political, lensink2000capital, boutchkova2012precarious, lehkonen2015democracy}. However, if market returns reflect investor expectations about future economic growth and expected returns, such hypotheses may be overly simplistic. Investors may view even a coup as a ``positive'' event if the current regime is anti-business or anti-global.

An active debate exists on the subject of ``good coups''---i.e., coups that lead to democratization or economic liberalization---and their effect on economic growth.  Most research suggests that ``good coups'' are not the norm \citep{derpanopoulos2015coups, powell2011global, thyne2016coup, varol2011democratic}, and others argue that the negative effects of uncertainty dominate the positive effects of coups staged by pro growth factions \citep{alesina1996political}, or that the expected impact of coups on economic growth is neutral as some enhance growth and others depress it \citep{londregan1990poverty}. But while ``good coups'' may be the minority, they have become more frequent \citep{marinov2014coups} and could have positive economic effects \citep{meyersson2016political, girardi2018institution}. Previous research therefore suggests that on average coups should not be expected to lead to increased economic development or positive market returns, but that positive returns associated with ``good coups'' could occasionally be present. As our event study approach also allows us to individual coups separately from others, we can evaluate this prediction.

Not all regime changes are equivalent---resignations, assassinations, and coups are all examples of sudden regime changes and have been used as proxies for instability in previous research. Some regime changes---such as the sudden resignation of an ineffective leader---may even lead to increased stability. By contrast, assassinations can occur seemingly at random, and a successor may be unclear. Coups can occur in the name of democracy or autocracy, and in general signal relative weakness in the current political system. These differences are often overlooked in past empirical strategies, which have proxied for instability in terms of the number of coups \citep{londregan1990poverty, alesina1996political}, assassinations or revolutions \citep{barro1991economic}, or have combined various metrics into single indices \citep{alesina1996income, venieris1986income, gupta1990economics, jong2009measurement}.\footnote{Note that these studies vary by metrics included in the indices, method of aggregation, and outcome variable of interest.} By contrast, we estimate effects separately for coups, assassinations, resignations, and protests.

Different types of regime changes may therefore have disparate effects on markets and/or economic growth. Taken together, the theory and research above suggests that coups should on average lead to negative returns, but sometimes lead to positive returns when the coup's instigators are clearly more pro market than the regime they replace. Assassinations should have a neutral or negative impact as they increase uncertainty, but institutional responses to assassinations vary by country. By contrast, resignations may be viewed positively on average, as they typically signal the departure of an ineffective leader.

Our empirical findings validate these predictions. We find that the type of political event and its expected impact on economic policy determines the direction of abnormal returns. Events expected to lead to more stable governance, economic liberalization, or democratization (such as willful resignations and coups that overthrow protectionist or leftist autocrats) are associated with positive returns, while those that consolidate authoritarian rule (e.g. military coups), exacerbate poor economic policies, or merely increase policy uncertainty (such as assassinations) have the opposite effect.

@@

Political data are primarily drawn from the Center for Systemic Peace's (CSP) Polity IV Coup d'etat dataset and Coup d'etat Events handbook. The Coup d'etat dataset includes the date of 1) successful coups, 2) attempted coups, 3) plotted coups and 4) alleged coup plots. We focus on successful coups because it is difficult to classify failed coups, and the choice of dataset and classification methodology would directly alter our findings.\footnote{\citet[p. 617]{needler1966political} has gone so far as to say that ``the categories of coups that were aborted, suppressed, or abandoned melt into each other and into a host of other non-coup phenomena so as to defy accounting,'' the Center for Systemic Peace states that it is ``confident that [its] list of successful coups is comprehensive'' but does not extend this confidence to attempted or failed coups, and \citet{powell2011global} state that it is ``difficult to identify more ambiguous forms of coup activity, such as coup failures, plots, and rumors.''}

@@

A list of the political events in our dataset is shown in \autoref{tab:regime-changes}. Coups tend to have the largest impact on the level of democratization as a number of countries have subsequently transitioned from democracies to anocracies or autocracies. On the other hand, neither assassinations or resignations have much of an impact on the level of democratization.

The discussion above shows that there is considerable debate about classification of regime changes. We recognize that some readers may feel that certain events are missing based on their own substantive knowledge. We rely on commonly used third-party classifications in order to minimize the possibility that our results are driven by our own subjective classifications. The one exception is that we separately analyze the failed coup attempt in Venezuela in April 2002 in which President Hugo Chavez was removed from office for two days, as it provides a natural test of the impact of the seemingly successful removal of a left-wing populist with a pro-business regime, and the ensuing reinstatement of a left-wing populist.

\autoref{fig:AV-DR} plots the absolute value of daily stock returns averaged across all events and show that the absolute value of returns on the event day (day 0) are significantly larger than on any other day. In addition, the magnitude of returns begins increasing just before the event day and remains high for a short period after. This suggests that financial volatility increases during the days surrounding regime changes, although it does not provide any evidence on mean returns. \nameref{sec: Impact of Political Instability on Stock Returns} tests these results more formally. It will first formally estimate the amount of volatility surrounding all irregular regime changes, then analyze coups, assassinations and resignations separately and determine what impact they have on mean returns.

\begin{figure}[!ht]
\includegraphics[width = 17cm]{../figs/daily_mean_absreturn.pdf}
\caption{Absolute value of daily returns}
\label{fig:AV-DR}
\end{figure}

@@

\autoref{fig:volatility} shows the mean volatility ($\overline{\sigma_t}$) estimates from the GARCH (1,1) model across all irregular regime changes for 250 trading days prior to and 250 days after each event. As expected, the volatility estimates stay between a narrow range at nearly all dates except those surrounding the regime change. Volatility appears to increase slowly just before the regime change, albeit not to a degree out of line with previous fluctuations in volatility. This may suggest that investors sometimes have information about the events before they occur. Nonetheless, there is still an enormous volatility jump on the day of the regime change. Volatility then decreases to normal levels within a month of the event.

@@

\subsection{Abnormal Returns} \label{subsec:abnormal-returns}

We now turn to the magnitude and direction of the effect of irregular regime changes on stock returns. We follow the standard event study methodology as presented by, among others, \citet{mackinlay1997event} and \citet{campbell1997econometrics}. Normal performance is measured with a constant mean return model,
\begin{align} \label{eqn:market-model}
R_{it}=\mu_{i}+\epsilon_{it},
\end{align}
where $R_{it}$ is the logged return of national stock index $i$ on trading day $t$ and $\epsilon_{it}$ is the error term. We calculate abnormal returns (ARs), in an ``event window'' surrounding the date of each coup, $AR_{i\tau}=R_{i\tau}-\widehat{\mu}_i$, where $\tau$ is a date in the event window, and $\widehat{\mu}_i$ is estimated in an ``estimation window'' preceding the event window with \autoref{eqn:market-model}. We use a 41 day event window (i.e. 20 pre-event trading days, the event day, and 20 post-event trading days). The estimation window is the 250 trading days prior to the start of the event window. The abnormal returns are then used to generate cumulative abnormal returns (CARs) between event day $\tau_1$ and event day $\tau_2$: $CAR(\tau_1,\tau_2)=\sum_{\tau=\tau_1}^{\tau_2}AR_{i\tau}$.

A constant mean return model is used instead of a market model in order to maximize the number of observations and because our unit of analysis is country-wide market indices rather than firms.\footnote{Plausible market indices such as the MSCI World Index and the S\&P/IFC Emerging Markets Investable Composite Index only begin in 1976 and 1995, respectively.} To address concerns regarding use of a constant mean return model, we also create a synthetic control portfolio for each event and compare observed returns to the synthetic returns.

We define the event date as the first trading day in which the market could have reacted to news of the event. For example, during the October 12, 1999 coup d'etat in Pakistan led by General Pervez Musharraf, the army announced that Prime Minister Nawaz Sharif had been dismissed after market hours at 10:15 pm. We code October 14th, the day in which the market re-opened, as the event day. When events occurred on weekends, we change the event date to the following Monday.

$(0,\tau-1)$ is used to denote the $\tau$-day period beginning with the event day and $(-1,\tau)$ to denote the negative $\tau$-day period beginning with the day prior to the event day. In other words, for cumulative abnormal returns prior to the event date, we aggregate backwards starting at the day of the event. For example, $CAR(-1,-2$) is the sum of the abnormal returns on event date $-1$ and event day $-2$. We present results for the sum of abnormal returns over the post-event windows of the event date only $(0,0$), the event date plus 6 days $(0,6$), and the event date plus 19 days $(0,19$). In addition, we present results for the pre-event windows $(-1,-7$) and $(-1,-20$).

As we hypothesize that different types of regime changes will have disparate effects on markets, we report abnormal returns separately for coups, assassinations and resignations. Standard errors and p-values are calculated using asymptotic t-statistics as in \citet{mackinlay1997event}.\footnote{It is appropriate to use the standard normal distribution to calculate test statistics because the length of the estimation window is sufficiently long (250 trading days).}

\subsection{Coups} \label{subsec: Coups}
\autoref{tab:AR-coups} shows abnormal returns for national stock indices both preceding and following coup d'etat. \autoref{tab:AR-coups} contains all coups presented in \autoref{tab:regime-changes} with the exception of the Argentinian coup of March 24, 1976. The March 24, 1976 Argentinian coup is excluded from our analysis because the stock market remained closed from March 24 to April 5, 1976, or a period of twelve days.\footnote{Treating this twelve day period as a single day CARs results in a positive abnormal return of 58\%, a fluctuation that seems qualitatively unreasonable.}

\begin{table}[!ht]
\caption{Abnormal returns following coups} \label{tab:AR-coups}
\vspace{-5pt}
\footnotesize
\begin{center}
\begin{threeparttable}
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}ld{4}d{4}d{4}d{4}d{4}d{2}}
  \hline
  \hline
\multicolumn{2}{c}{} & \multicolumn{3}{c}{Post-Event CAR} & \multicolumn{2}{c}{Pre-Event CAR} & \multicolumn{1}{c}{\multirow{2}{*}{Days to}}\\
\cmidrule(r){3-5} \cmidrule(r){6-7}
\multicolumn{1}{c}{Country} & \multicolumn{1}{c}{Event Date} & \multicolumn{1}{c}{(0,0)} & \multicolumn{1}{c}{(0,6)} & \multicolumn{1}{c}{(0,19)} & \multicolumn{1}{c}{(-1,-7)} & \multicolumn{1}{c}{(-1,-20)} & \multicolumn{1}{c}{rebound}\\
  \hline
\ExpandableInput{../tables/artable-coups-car.txt}
  \hline
\ExpandableInput{../tables/artable-coups-car-mean.txt}
   \hline
   \hline
\end{tabular*}
\scriptsize
Notes: Standard errors are in parentheses. ``Days to rebound'' is the number of trading days following a negative stock return for the national stock index to return to pre-event level (it is calculated if the price decreases on the event day, not if the event day abnormal return is negative). Returns are inflation adjusted. 
\end{threeparttable}
\end{center}
\end{table}

@@

The results in \autoref{tab:AR-ass} are produced from analyses identical to those in \autoref{tab:AR-coups} but for successful assassinations rather than coups. Like the majority of coups, there is evidence that assassinations decrease stock prices. The mean event day abnormal return is negative and statistically different than zero. However, the result is driven by five events: the shooting of U.S. President William McKinley on September 6, 1901; the assassination of U.S. President John F. Kennedy on November 22, 1963; the assassination of Indian Prime Minister Indira Gandhi on October 31, 1984; the suicide bombing that killed Sri Lankan president Ranasinghe Premadasa on May 1, 1993; and the assassination of Israeli Prime Minister Yitzhak Rabin on the evening of November 4, 1995.

\begin{table}[!ht]
\caption{Abnormal returns following assassinations} \label{tab:AR-ass}
\vspace{-5pt}
\footnotesize
\begin{center}
\begin{threeparttable}
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}ld{4}d{4}d{4}d{4}d{4}d{2}}
  \hline
  \hline
\multicolumn{2}{c}{} & \multicolumn{3}{c}{Post-Event CAR} & \multicolumn{2}{c}{Pre-Event CAR} & \multicolumn{1}{c}{\multirow{2}{*}{Days to}}\\
\cmidrule(r){3-5} \cmidrule(r){6-7}
\multicolumn{1}{c}{Country} & \multicolumn{1}{c}{Event Date} & \multicolumn{1}{c}{(0,0)} & \multicolumn{1}{c}{(0,6)} & \multicolumn{1}{c}{(0,19)} & \multicolumn{1}{c}{(-1,-7)} & \multicolumn{1}{c}{(-1,-20)} & \multicolumn{1}{c}{rebound}\\
  \hline
\ExpandableInput{../tables/artable-ass-car.txt}
  \hline
\ExpandableInput{../tables/artable-ass-car-mean.txt}
   \hline
   \hline
\end{tabular*}
\scriptsize
Notes: Standard errors are in parentheses. ``Days to rebound'' is the number of trading days following a negative stock return for the national stock index to return to pre-event level (it is calculated if the price decreases on the event day, not if the event day abnormal return is negative). Returns are inflation adjusted. 
\end{threeparttable}
\end{center}
\end{table}

These results are consistent with our hypothesis that the nature of the political event and its expected impact on policy matters, and that assassinations should have a negative effect as they occur seemingly at random and increase uncertainty. While the mean effect of assassinations is negative, it is smaller in magnitude than for coups. Unlike a coup, an assassination may not necessarily be expected to cause immediate change in economic policy, particularly in the presence of an institutionalized line of succession. As such, we would expect CARs to be negative due to increased instability and uncertainty, but smaller in magnitude to a coup or resignation due to greater expectations of policy inertia.

@@

These results are again consistent with our hypothesis that different events will have disparate effects, and that resignations may lead to positive returns. The positive event day abnormal return following resignations is not surprising as resignations typically occur because of poor performance and/or loss of authority. Among our sample of events, leaders were ousted following loss in conflict/war, anti-authoritarian protest, corruption scandals, Supreme Court ruling against unconstitutional actions, contested elections, and abuse of power.

For example, consider Ferdinand Marcos' resignation from office as President of the Philippines in February 1986. Prior to his resignation, the Philippine regime was known for rampant corruption, crony capitalism, extreme inequality, high unemployment, failed import substitution industrialization policy, and oligarchic control of the economy \citep{overholt1986rise, traywick2014}. In fact, the Philippines was the least preferred site for foreign investment amongst East Asian capitalist economies and possessed one of the worst capital investment to economic output ratios in Asia \citep{overholt1986rise}. Marcos held a snap presidential election on February 7, 1986, in which he declared victory despite overwhelming evidence of electoral fraud. Public protests ensued, and two weeks later the military withdrew its support of the Marcos regime \citep{lee2009armed}. Marcos was replaced by his electoral opponent, Corazon Aquino, who had run on a platform of economic liberalization and elimination of crony capitalism \citep{villegas1987philippines}. This event was associated with an approximately 13\% positive event day AR.

By contrast, the largest negative event in our sample (-3\%) is the 1993 resignation of President Ghulam Ishaq Khan and Prime Minister Nawaz Sharif in Pakistan. The resignations occurred after months of political infighting when the army demanded the President and Prime Minister resign and call for new elections. An interim prime minister was installed, but uncertainty about Pakistan's political and economic future remained high prior to the next round of elections.

@@

A similar analysis is not possible for shifts from pro or anti business leaders, as examples of such clear and plausibly exogenous shifts in leader economic ideology do not exist in our sample.\footnote{Based on matching our cases with codings from the The Ideology of Heads of Government (HOG) database, as well as surveys of news reports on the day of each event. In all cases, no clear economic ideological shift can be identified.} We therefore look outside of the sample above and conduct an in-depth case study of a seemingly pro-business and anti-socialist coup followed by the reinstatement of a socialist government: the 2002 failed coup against Hugo Chavez in Venezuela.

@@

The ultimately failed Venezuelan coup against Hugo Chavez replaced a left-wing populist government with a new pro-business president, and therefore provides a natural test of the effects of both pro-business and anti-business regime changes separately from simple uncertainty because investors reacted to an expected regime change twice: first, when Chavez was ousted, and second, when he was reinstated. On the evening of April 11, 2002, coup plotters removed Chavez from office and later detained him. Pedro Carmona, a Venezuelan economist and business leader, was named the transitional President of Venezuela. Two days later, on April 13, 2002, a popular uprising led to Chavez's reinstatement as president. This provides an estimate of the market's valuation of a transition from the Chavez regime to the Carmona regime and its valuation of a transition from the Carmona regime back to the Chavez regime. By extension, it provides an estimate of the impact of a shift from a left-wing populist government to a pro business regime in an emerging market.

\autoref{fig:AR-Ven} provides graphical evidence on the effect of the coup attempt. The top panel shows CARs for the 10 days prior to and following the event, along with 95\% confidence intervals. The daily ARs and corresponding confidence intervals are displayed in the bottom panel. The abnormal return on April 12, the first trading day in which investors could react to the coup, was +10\%. The market reacted in the opposite direction to Chavez's reinstatement as president: the abnormal return on the next trading day, April 15, 2002, was -8\%.

>>>>

\section{Reexamining market response to irregular regime changes}

Conventional theory and cross-country empirical evidence suggests that political instability depresses financial returns and economic growth \citep{irshad2017relationship, le2006political, lensink2000capital, boutchkova2012precarious, lehkonen2015democracy}. However, if market returns reflect investor expectations about future economic growth and returns, such hypotheses may be too simplistic. Investors may view even a coup as a ``positive'' event if the current regime is anti-business or anti-global.

%Previous research therefore suggests that coups should on average not lead to positive economic development or market returns, but that ``good coups'' could occasionally create positive returns. As our event study approach allows us to analyze individual coups separately from others, we can test this prediction.

Not all regime changes are equivalent. Resignations, assassinations, and coups have all been used as proxies for instability in previous research. However, some regime changes---such as the sudden resignation of an ineffective leader---may lead to increased stability. By contrast, assassinations can occur seemingly at random and a successor may be unclear. Coups can occur in the name of democracy or autocracy. These differences are often overlooked in past empirical strategies, which have proxied for instability in terms of the number of coups \citep{londregan1990poverty, alesina1996political}, assassinations or revolutions \citep{barro1991economic}, or have combined various metrics into single indices \citep{alesina1996income, venieris1986income, gupta1990economics, jong2009measurement}.\footnote{Note that these studies vary by metrics included in the indices, method of aggregation, and outcome variable of interest.} By contrast, we estimate effects separately for coups, assassinations, resignations, and protests.

An active debate also exists on the subject of ``good coups''---i.e., coups that lead to democratization or economic liberalization---and their effect on economic growth.  Most research suggests that ``good coups'' are not the norm \citep{derpanopoulos2015coups, powell2011global, thyne2016coup, varol2011democratic}, others argue that the negative effects of uncertainty dominate any positive effects \citep{alesina1996political}, or that the expected impact of coups  is neutral as some enhance growth while others depress it \citep{londregan1990poverty}. But while ``good coups'' may be the minority, they have become more frequent \citep{marinov2014coups} and could have positive economic effects \citep{meyersson2016political, girardi2018institution}.

Different types of regime changes may therefore have disparate effects on markets and/or economic growth. The theory and research above suggests that coups should on average lead to negative returns, but sometimes lead to positive returns when the coup's instigators are clearly more pro-market than the regime they replace. As our event study approach allows us to analyze individual coups separately from others, we can test this prediction. Assassinations should have a neutral or negative impact as they increase uncertainty, but institutional responses to assassinations vary by country. By contrast, resignations may be viewed positively on average, as they typically signal the departure of an ineffective leader.

%Our empirical findings validate these predictions. We find that the type of political event and its expected impact on economic policy determines the direction of abnormal returns. Events expected to lead to more stable governance, economic liberalization, or democratization (such as willful resignations and coups that overthrow protectionist or leftist autocrats) are associated with positive returns, while those that consolidate authoritarian rule (e.g. military coups), exacerbate poor economic policies, or merely increase policy uncertainty (such as assassinations) have the opposite effect.

@@

Political data are primarily drawn from the Center for Systemic Peace's (CSP) Polity IV Coup d'etat dataset and Coup d'etat Events handbook. The Coup d'etat dataset includes the date of successful, attempted, plotted, and alleged coups. We focus on successful coups because the other classifications are contentious, and the choice of dataset and classification methodology would directly alter our findings.\footnote{\citet[p. 617]{needler1966political} has gone so far as to say that ``the categories of coups that were aborted, suppressed, or abandoned melt into each other and into a host of other non-coup phenomena so as to defy accounting,'' the Center for Systemic Peace states that it is ``confident that [its] list of successful coups is comprehensive'' but does not extend this confidence to attempted or failed coups, and \citet{powell2011global} state that it is ``difficult to identify more ambiguous forms of coup activity, such as coup failures, plots, and rumors.''}

@@

A list of the political events in our dataset and whether they resulted in democratic or autocratic shifts is shown in \autoref{tab:regime-changes}. Coups tend to have the largest impact on the level of democratization as a number of countries have subsequently transitioned from democracies to anocracies or autocracies. On the other hand, neither assassinations nor resignations tend to have much impact on the level of democratization.

The discussion above shows that there is considerable debate about classification of regime changes. We recognize that some readers may feel certain events are missing based on their own substantive knowledge. We rely on commonly used third-party classifications in order to minimize the possibility that our results are driven by our own subjective classifications. The one exception is that we separately analyze the failed coup attempt in Venezuela in April 2002 in which President Hugo Chavez was removed from office for two days, as it provides a natural test of the impact of the seemingly successful removal of a left-wing populist with a pro-business regime, and the ensuing reinstatement of a left-wing populist.

\autoref{fig:AV-DR} plots the absolute value of daily stock returns averaged across all events and shows that the absolute value of returns on the event day (day 0) are significantly larger than on any other day. In addition, the magnitude of returns begins increasing just before the event day and remains high for a short period after. This suggests that financial volatility increases during the days surrounding regime changes, although it does not provide any evidence on mean returns. \nameref{sec: Impact of Political Instability on Stock Returns} tests these results more formally. It will first formally estimate the amount of volatility surrounding all irregular regime changes, then analyze coups, assassinations and resignations separately and determine what impact they have on mean returns.

\begin{figure}[!ht]
\begin{centering}
\includegraphics[width = 15cm]{../figs/daily_mean_absreturn.pdf}
\caption{Absolute value of daily returns}
\label{fig:AV-DR}
\end{centering}
\end{figure}

@@

\autoref{fig:volatility} shows the mean volatility ($\overline{\sigma_t}$) estimates across all irregular regime changes for 250 trading days prior to and 250 days after each event. Volatility stays between a narrow range at nearly all dates except those surrounding the regime change. Volatility appears to increase slowly just before the regime change, albeit not to a degree out of line with previous fluctuations. This may suggest that investors sometimes have information about the events before they occur. Nonetheless, there is still an enormous volatility jump on the day of the regime change. Volatility then decreases to normal levels within a month of the event.

@@

\subsection{Estimation strategy: abnormal returns} \label{subsec:abnormal-returns}

We now turn to the magnitude and direction of the effect of irregular regime changes on stock returns. We follow the standard event study methodology as presented by, among others, \citet{mackinlay1997event} and \citet{campbell1997econometrics}. Normal performance is measured with a constant mean return model,
\begin{align} \label{eqn:market-model}
R_{it}=\mu_{i}+\epsilon_{it},
\end{align}
where $R_{it}$ is the logged return of national stock index $i$ on trading day $t$ and $\epsilon_{it}$ is the error term. We calculate abnormal returns (ARs), in an ``event window'' surrounding the date of each coup, $AR_{i\tau}=R_{i\tau}-\widehat{\mu}_i$, where $\tau$ is a date in the event window, and $\widehat{\mu}_i$ is estimated in an ``estimation window'' preceding the event window with \autoref{eqn:market-model}. We use a 41 day event window (i.e. 20 pre-event trading days, the event day, and 20 post-event trading days). The estimation window is the 250 trading days prior to the start of the event window. The abnormal returns are then used to generate cumulative abnormal returns (CARs) between event day $\tau_1$ and event day $\tau_2$: $CAR(\tau_1,\tau_2)=\sum_{\tau=\tau_1}^{\tau_2}AR_{i\tau}$. Standard errors and p-values are calculated using asymptotic t-statistics as in \citet{mackinlay1997event}.\footnote{It is appropriate to use the standard normal distribution to calculate test statistics because the length of the estimation window is sufficiently long (250 trading days).}

We define the event date as the first trading day in which the market could have reacted to news of the event.\footnote{For example, during the October 12, 1999 coup d'etat in Pakistan led by General Pervez Musharraf, the army announced that Prime Minister Nawaz Sharif had been dismissed after market hours at 10:15 pm. We code October 14th, the day in which the market re-opened, as the event day.} When events occurred on weekends, we change the event date to the following Monday. $(0,\tau-1)$ is used to denote the $\tau$-day period beginning with the event day and $(-1,\tau)$ to denote the negative $\tau$-day period beginning with the day prior to the event day.\footnote{In other words, for cumulative abnormal returns prior to the event date, we aggregate backwards starting at the day of the event. For example, $CAR(-1,-2$) is the sum of the abnormal returns on event date $-1$ and event day $-2$.} We present results for the sum of abnormal returns over the post-event windows of the event date only $(0,0$), the event date plus 6 days $(0,6$), and the event date plus 19 days $(0,19$). In addition, we present results for the pre-event windows $(-1,-7$) and $(-1,-20$).

We use a constant mean return model instead of a market model to maximize the number of observations and because our unit of analysis \textit{is} the country-wide market index (not firms).\footnote{Plausible market indices such as the MSCI World Index and the S\&P/IFC Emerging Markets Investable Composite Index only begin in 1976 and 1995, respectively.} To address concerns regarding use of a constant mean return model, we integrate synthetic control methodology \citep{abadie2003economic, abadie2010synthetic} with event study estimation. In short, we create a ``synthetic'' control portfolio for each event and compare observed returns to the synthetic returns. Each country is given a weight which represents its influence in the synthetic control portfolio. The weight is chosen so that the daily returns and the variance of the daily returns of the control portfolio and the event country are most similar in the estimation window.

More formally, let $\boldsymbol{R}_{k}$ be the vector of returns for the event country in the estimation window, $\boldsymbol{R}_{-k}$ be the vector of returns for all other countries in the estimation window, $\boldsymbol{X}_1=(\boldsymbol{R}_{k},\var(\boldsymbol{R}_{k}))$, $\boldsymbol{X}_0=(\boldsymbol{R}_{-k},\var(\boldsymbol{R}_{-k}))$, and $\boldsymbol{W}_{-k}$ be a $((N-1) \times 1)$ vector of weights where $N$ is the number of countries listed in \autoref{tab:stock-list}. Then $\boldsymbol{W}^*$ is chosen to minimize $(\boldsymbol{X}_1-\boldsymbol{X}_0\boldsymbol{W})'\boldsymbol{V}(\boldsymbol{X}_1-\boldsymbol{X}_0\boldsymbol{W})$ subject to $w_i\geq0$ $(i = 1,2,\ldots,N-1)$ and $\sum_i^{N-1} w_i = 1$, and the vector $\boldsymbol{V}$ is chosen so that stock returns for the control portfolio during the estimation window are are close as possible to the event country.\footnote{See \citet{abadie2003economic} for further details.} The set of possible countries in the control portfolio consists of all countries listed in \autoref{tab:stock-list}, and the results are presented in \nameref{subsec: Robustness}.

\subsection{Coups} \label{subsec: Coups}

\autoref{tab:AR-coups} shows abnormal returns for national stock indices both preceding and following coup d'etat. \autoref{tab:AR-coups} contains all coups presented in \autoref{tab:regime-changes}, with the exception of the Argentinian coup of March 24, 1976 which is excluded from our analysis because the stock market remained closed for twelve days following the event.\footnote{Treating this twelve day period as a single day CARs results in a positive abnormal return of 58\%, a fluctuation that seems qualitatively unreasonable.}

\begin{table}[!htb]
\caption{Abnormal returns following coups} \label{tab:AR-coups}
\vspace{-5pt}
\footnotesize
\begin{center}
\begin{threeparttable}
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}ld{4}d{4}d{4}d{4}d{4}d{2}}
  \hline
  \hline
\multicolumn{2}{c}{} & \multicolumn{3}{c}{Post-Event CAR} & \multicolumn{2}{c}{Pre-Event CAR} & \multicolumn{1}{c}{\multirow{2}{*}{Days to}}\\
\cmidrule(r){3-5} \cmidrule(r){6-7}
\multicolumn{1}{c}{Country} & \multicolumn{1}{c}{Event Date} & \multicolumn{1}{c}{(0,0)} & \multicolumn{1}{c}{(0,6)} & \multicolumn{1}{c}{(0,19)} & \multicolumn{1}{c}{(-1,-7)} & \multicolumn{1}{c}{(-1,-20)} & \multicolumn{1}{c}{rebound}\\
  \hline
\ExpandableInput{../tables/artable-coups-car.txt}
  \hline
\ExpandableInput{../tables/artable-coups-car-mean.txt}
   \hline
   \hline
\end{tabular*}
\scriptsize
Notes: Standard errors are in parentheses. ``Days to rebound'' is the number of trading days following a negative stock return for the national stock index to return to pre-event level (it is calculated if the price decreases on the event day, not if the event day abnormal return is negative). Returns are inflation adjusted. 
\end{threeparttable}
\end{center}
\end{table}

@@

The results in \autoref{tab:AR-ass} are produced from analyses identical to those in \autoref{tab:AR-coups} but for successful assassinations rather than coups. Like the majority of coups, there is evidence that assassinations decrease stock prices. The mean event day abnormal return is negative and statistically different than zero.

%However, the result is driven by five events: the shooting of U.S. President William McKinley on September 6, 1901; the assassination of U.S. President John F. Kennedy on November 22, 1963; the assassination of Indian Prime Minister Indira Gandhi on October 31, 1984; the suicide bombing that killed Sri Lankan president Ranasinghe Premadasa on May 1, 1993; and the assassination of Israeli Prime Minister Yitzhak Rabin on the evening of November 4, 1995.

\begin{table}[!htb]
\caption{Abnormal returns following assassinations} \label{tab:AR-ass}
\vspace{-5pt}
\footnotesize
\begin{center}
\begin{threeparttable}
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}ld{4}d{4}d{4}d{4}d{4}d{2}}
  \hline
  \hline
\multicolumn{2}{c}{} & \multicolumn{3}{c}{Post-Event CAR} & \multicolumn{2}{c}{Pre-Event CAR} & \multicolumn{1}{c}{\multirow{2}{*}{Days to}}\\
\cmidrule(r){3-5} \cmidrule(r){6-7}
\multicolumn{1}{c}{Country} & \multicolumn{1}{c}{Event Date} & \multicolumn{1}{c}{(0,0)} & \multicolumn{1}{c}{(0,6)} & \multicolumn{1}{c}{(0,19)} & \multicolumn{1}{c}{(-1,-7)} & \multicolumn{1}{c}{(-1,-20)} & \multicolumn{1}{c}{rebound}\\
  \hline
\ExpandableInput{../tables/artable-ass-car.txt}
  \hline
\ExpandableInput{../tables/artable-ass-car-mean.txt}
   \hline
   \hline
\end{tabular*}
\scriptsize
Notes: Standard errors are in parentheses. ``Days to rebound'' is the number of trading days following a negative stock return for the national stock index to return to pre-event level (it is calculated if the price decreases on the event day, not if the event day abnormal return is negative). Returns are inflation adjusted. 
\end{threeparttable}
\end{center}
\end{table}

These results are consistent with our hypothesis that assassinations should have a negative effect as they occur seemingly at random and increase uncertainty. While the mean effect of assassinations is negative, it is smaller in magnitude than for coups. Unlike a coup, an assassination may not necessarily be expected to cause immediate change in economic policy, particularly in the presence of an institutionalized line of succession. As such, we would expect CARs to be negative due to increased instability and uncertainty, but smaller in magnitude to a coup or resignation due to greater expectations of policy inertia.

@@

These results are again consistent with our hypothesis that different events will have disparate effects, and that resignations may lead to positive returns as resignations typically occur because of poor performance and/or loss of authority. Among our sample of events, leaders were ousted following loss in conflict/war, anti-authoritarian protest, corruption scandals, Supreme Court ruling against unconstitutional actions, contested elections, and abuse of power.

For example, the resignation of Ferdinand Marcos as President of the Philippines in February 1986 was associated with an approximately 13\% positive event day AR. Prior to his resignation, the Philippine regime was known for rampant corruption, crony capitalism, extreme inequality, high unemployment, failed import substitution industrialization policy, and oligarchic control of the economy \citep{overholt1986rise, traywick2014}. In fact, the Philippines was the least preferred site for foreign investment amongst East Asian capitalist economies \citep{overholt1986rise}.

%For example, consider Ferdinand Marcos' resignation from office as President of the Philippines in February 1986. Prior to his resignation, the Philippine regime was known for rampant corruption, crony capitalism, extreme inequality, high unemployment, failed import substitution industrialization policy, and oligarchic control of the economy \citep{overholt1986rise, traywick2014}. In fact, the Philippines was the least preferred site for foreign investment amongst East Asian capitalist economies and possessed one of the worst capital investment to economic output ratios in Asia \citep{overholt1986rise}. Marcos held a snap presidential election on February 7, 1986, in which he declared victory despite overwhelming evidence of electoral fraud. Public protests ensued, and two weeks later the military withdrew its support of the Marcos regime \citep{lee2009armed}. Marcos was replaced by his electoral opponent, Corazon Aquino, who had run on a platform of economic liberalization and elimination of crony capitalism \citep{villegas1987philippines}. This event was associated with an approximately 13\% positive event day AR.

%By contrast, the largest negative event in our sample (-3\%) is the 1993 resignation of President Ghulam Ishaq Khan and Prime Minister Nawaz Sharif in Pakistan. The resignations occurred after months of political infighting when the army demanded the President and Prime Minister resign and call for new elections. An interim prime minister was installed, but uncertainty about Pakistan's political and economic future remained high prior to the next round of elections.

@@

A similar analysis is not possible for shifts from pro or anti business leaders, as examples of such clear and plausibly exogenous shifts in leader economic ideology do not exist in our sample.\footnote{Based on matching our cases with codings from the The Ideology of Heads of Government (HOG) database, as well as surveys of news reports on the day of each event. In all cases, no clear economic ideological shift can be identified.} We therefore look outside of the sample above and conduct an in-depth case study of a seemingly pro-business and anti-socialist coup followed by the reinstatement of a socialist government: the 2002 failed coup against Hugo Chavez in Venezuela. The coup replaced a left-wing populist government with a new pro-business president, and therefore provides a natural test of the effects of both pro-business and anti-business regime changes separately from simple uncertainty because investors reacted to an expected regime change twice: first, when Chavez was ousted, and second, when he was reinstated.

@@

%The ultimately failed Venezuelan coup against Hugo Chavez replaced a left-wing populist government with a new pro-business president, and therefore provides a natural test of the effects of both pro-business and anti-business regime changes separately from simple uncertainty because investors reacted to an expected regime change twice: first, when Chavez was ousted, and second, when he was reinstated. On the evening of April 11, 2002, coup plotters removed Chavez from office and later detained him. Pedro Carmona, a Venezuelan economist and business leader, was named the transitional President of Venezuela. Two days later, on April 13, 2002, a popular uprising led to Chavez's reinstatement as president. This provides an estimate of the market's valuation of a transition from the Chavez regime to the Carmona regime and its valuation of a transition from the Carmona regime back to the Chavez regime. By extension, it provides an estimate of the impact of a shift from a left-wing populist government to a pro business regime in an emerging market.
